---
title: "COMPSCIX 415.2 Homework 1"
author: "Kannu-priya"
date: "February 4, 2018"
output: html_document
---

##Set up


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (mdsr)
library (tidyverse)
```


My Github repository for my assignments can be found at this URL: https://github.com/kannu-priya/compscix-415-2-assignments.git

#Explore WorldCities data frame

About World Cities Data:
The Data set has list of 23018 cities as rows (observations) and 10 variable columns covering some basic attributes about the cities. The variables in the data set are - City Name, 2 letter country code, country region, ISO city code, latitude & longitude, population etc. 

Lets dig into WorldCities data set


```{r load_package, warning=FALSE, message=FALSE}
data("WorldCities")
glimpse(WorldCities)
WorldCities <- head(WorldCities, 200)
country_col <- WorldCities$country
unique(country_col)
region <- WorldCities$region
unique(region)
WorldCities %>% select(country) %>% unique()
WorldCities %>% ggplot (aes (x=population)) +
  geom_histogram() +
  xlab ('City Population')

WorldCities %>% ggplot (aes (x=population)) +
  geom_histogram() +
  xlab ('City Total Population') +
  ylab ('Count of Cities')+
  ggtitle ('Population Distribution Chart')+
  theme_bw()


```